---
title: "seasonal trends"
output: 
  html_notebook:
    toc: true
    code-fold: true
knitr: true
---

install libraries
```{r}
#| code-fold: true
shh <- suppressMessages
shh(library(DBI))
shh(library(RPostgres))
shh(library(tidyverse))
shh(library(lubridate))
shh(library(plotly))
shh(library(glue))
```

set up connection to postgres
```{r}
#| code-fold: true
connection <- dbConnect(
    Postgres(),
    dbname = "metalworkingsolutions",
    host = "localhost",
    port = 5432,
    user = "postgres",
    password = "postgres"
    # password = rstudioapi::askForPassword("Database Password")
)
```

create query for top 20 customers for seasonal evaluation
```{r}
#| code-fold: true
query <- "
SELECT
    omp_customer_organization_id,
    omp_sales_order_id,
    omp_customer_po,
    omp_order_date,
    uomp_promise_date,
    omp_payment_term_id,
    omp_order_total_base
FROM sales_orders;
"
result_set <- dbSendQuery(connection, query)
```

create tibble for customers
```{r}
#| code-fold: true
customers <- dbFetch(result_set)
dbClearResult(result_set)
```


### get top 20 customers by generated revenue for 2023, 2024, and both years combined

2023
```{r}
#| code-fold: true
top20_customers_2023 <- customers |>
    select(omp_order_date, omp_customer_organization_id, omp_order_total_base) |>
    group_by(omp_customer_organization_id) |>
    mutate(omp_order_date = year(omp_order_date)) |>
    filter(omp_order_date %in% 2023) |>
    summarise(generated_revenue_2023 = sum(omp_order_total_base)) |>
    rename(customer_id = omp_customer_organization_id) |>
    arrange(desc(generated_revenue_2023)) |>
    select(customer_id) |>
    head(20)
```

2024
```{r}
#| code-fold: true
top20_customers_2024 <- customers |>
    select(omp_order_date, omp_customer_organization_id, omp_order_total_base) |>
    group_by(omp_customer_organization_id) |>
    mutate(omp_order_date = year(omp_order_date)) |>
    filter(omp_order_date %in% 2024) |>
    summarise(generated_revenue_2023 = sum(omp_order_total_base)) |>
    rename(customer_id = omp_customer_organization_id) |>
    arrange(desc(generated_revenue_2023)) |>
    select(customer_id) |>
    head(20)
```

2023 & 2024
```{r}
#| code-fold: true
top20_customers_total <- customers |>
    select(omp_order_date, omp_customer_organization_id, omp_order_total_base) |>
    group_by(omp_customer_organization_id) |>
    mutate(omp_order_date = year(omp_order_date)) |>
    summarise(generated_revenue_2023 = sum(omp_order_total_base)) |>
    rename(customer_id = omp_customer_organization_id) |>
    arrange(desc(generated_revenue_2023)) |>
    select(customer_id) |>
    head(20)
```


## monthly change by customer, 2023
Using plotly for interactivity. 
```{r, out.width='100%'}
#| code-fold: true
p <- customers |>
    select(omp_order_date, omp_customer_organization_id, omp_order_total_base) |>
    mutate(month = month(omp_order_date), year = year(omp_order_date)) |>
    rename(
        customer_id = omp_customer_organization_id,
        generated_revenue = omp_order_total_base
    ) |>
    filter(year == 2023, customer_id %in% top20_customers_2023$customer_id) |>
    group_by(customer_id, month) |>
    summarise(generated_revenue = sum(generated_revenue), .groups = "drop") |>
    mutate(
        customer_id = factor(customer_id,
            levels = top20_customers_2023$customer_id
        )
    ) |>
    ggplot(aes(x = month, y = generated_revenue, color = customer_id)) +
    geom_line() +
    scale_x_continuous(
        breaks = seq_along(month.name),
        labels = month.name
    ) +
    labs(
        title = "Change in Customer Generated Revenue 2023",
        x = "Month",
        y = "Revenue ($)"
    ) +
    theme_minimal() +
    theme(
        axis.text.x = element_text(angle = 55, hjust = 1)
    )
ggplotly(p)
```

ggplot version for displaying inline better.
```{r, fig.width=10}
#| code-fold: true
customers |>
    select(omp_order_date, omp_customer_organization_id, omp_order_total_base) |>
    mutate(month = month(omp_order_date), year = year(omp_order_date)) |>
    rename(
        customer_id = omp_customer_organization_id,
        generated_revenue = omp_order_total_base
    ) |>
    filter(year == 2023, customer_id %in% top20_customers_2023$customer_id) |>
    group_by(customer_id, month) |>
    summarise(generated_revenue = sum(generated_revenue), .groups = "drop") |>
    mutate(
        customer_id = factor(customer_id,
            levels = top20_customers_2023$customer_id
        )
    ) |>
    ggplot(aes(x = month, y = generated_revenue, color = customer_id)) +
    geom_line() +
    scale_x_continuous(
        breaks = seq_along(month.name),
        labels = month.name
    ) +
    labs(
        title = "Change in Customer Generated Revenue 2023",
        x = "Month",
        y = "Revenue ($)"
    ) +
    theme_minimal() +
    theme(
        axis.text.x = element_text(angle = 55, hjust = 1)
    )
```

## monthly change by customer, 2024
Using plotly for interactivity 
```{r}
#| code-fold: true
p <- customers |>
    select(omp_order_date, omp_customer_organization_id, omp_order_total_base) |>
    mutate(month = month(omp_order_date), year = year(omp_order_date)) |>
    rename(
        customer_id = omp_customer_organization_id,
        generated_revenue = omp_order_total_base
    ) |>
    filter(year == 2024, customer_id %in% top20_customers_2024$customer_id) |>
    group_by(customer_id, month) |>
    summarise(generated_revenue = sum(generated_revenue), .groups = "drop") |>
    mutate(
        customer_id = factor(customer_id,
            levels = top20_customers_2024$customer_id
        )
    ) |>
    ggplot(aes(x = month, y = generated_revenue, color = customer_id)) +
    geom_line() +
    scale_x_continuous(
        breaks = seq_along(month.name),
        labels = month.name
    ) +
    labs(
        title = "Change in Customer Generated Revenue 2024",
        x = "Month",
        y = "Revenue ($)"
    ) +
    theme_minimal() +
    theme(
        axis.text.x = element_text(angle = 55, hjust = 1)
    )
ggplotly(p)
```

using ggplot for better inline visibility
```{r, fig.width=10}
#| code-fold: true
customers |>
    select(omp_order_date, omp_customer_organization_id, omp_order_total_base) |>
    mutate(month = month(omp_order_date), year = year(omp_order_date)) |>
    rename(
        customer_id = omp_customer_organization_id,
        generated_revenue = omp_order_total_base
    ) |>
    filter(year == 2024, customer_id %in% top20_customers_2024$customer_id) |>
    group_by(customer_id, month) |>
    summarise(generated_revenue = sum(generated_revenue), .groups = "drop") |>
    mutate(
        customer_id = factor(customer_id,
            levels = top20_customers_2024$customer_id
        )
    ) |>
    ggplot(aes(x = month, y = generated_revenue, color = customer_id)) +
    geom_line() +
    scale_x_continuous(
        breaks = seq_along(month.name),
        labels = month.name
    ) +
    labs(
        title = "Change in Customer Generated Revenue 2024",
        x = "Month",
        y = "Revenue ($)"
    ) +
    theme_minimal() +
    theme(
        axis.text.x = element_text(angle = 55, hjust = 1)
    )
```

2023 + 2024
```{r}
#| code-fold: true
p <- customers |>
    select(omp_order_date, omp_customer_organization_id, omp_order_total_base) |>
    mutate(
        month = month(omp_order_date),
        year = year(omp_order_date),
        month_year = make_date(year, month)
    ) |>
    rename(
        customer_id = omp_customer_organization_id,
        generated_revenue = omp_order_total_base
    ) |>
    filter(customer_id %in% top20_customers_total$customer_id) |>
    group_by(customer_id, month_year) |> # Group by month_year
    summarise(generated_revenue = sum(generated_revenue), .groups = "drop") |>
    mutate(
        customer_id = factor(customer_id,
            levels = top20_customers_total$customer_id
        )
    ) |>
    ggplot(aes(x = month_year, y = generated_revenue, color = customer_id)) +
    geom_line() +
    scale_x_date(date_breaks = "1 month", date_labels = "%b %Y") + # Format x-axis
    labs(
        title = "Change in Customer Generated Revenue 2023 + 2024",
        x = "Month",
        y = "Revenue ($)"
    ) +
    theme_minimal() +
    theme(
        axis.text.x = element_text(angle = 55, hjust = 1)
    )

ggplotly(p)
```

### ggplot for inline visibility
```{r, fig.width=10}
#| code-fold: true
customers |>
    select(omp_order_date, omp_customer_organization_id, omp_order_total_base) |>
    mutate(
        month = month(omp_order_date),
        year = year(omp_order_date),
        month_year = make_date(year, month)
    ) |>
    rename(
        customer_id = omp_customer_organization_id,
        generated_revenue = omp_order_total_base
    ) |>
    filter(customer_id %in% top20_customers_total$customer_id) |>
    group_by(customer_id, month_year) |> # Group by month_year
    summarise(generated_revenue = sum(generated_revenue), .groups = "drop") |>
    mutate(
        customer_id = factor(customer_id,
            levels = top20_customers_total$customer_id
        )
    ) |>
    ggplot(aes(x = month_year, y = generated_revenue, color = customer_id)) +
    geom_line() +
    scale_x_date(date_breaks = "1 month", date_labels = "%b %Y") + # Format x-axis
    labs(
        title = "Change in Customer Generated Revenue 2023 + 2024",
        x = "Month",
        y = "Revenue ($)"
    ) +
    theme_minimal() +
    theme(
        axis.text.x = element_text(angle = 55, hjust = 1)
    )
```

### New query to select jobs
```{r}
#| code-fold: true
query <- "
SELECT
    jmp_customer_organization_id AS customer_id,
    jmp_job_id AS job_id,
    jmp_part_id AS part_id,
    jmp_part_short_description AS part_description,
    jmp_order_quantity AS order_quantity,
    jmp_production_quantity AS production_quantity,
    jmp_production_due_date AS production_due_date

FROM jobs;
"
result_set <- dbSendQuery(connection, query)
jobs <- dbFetch(result_set)
dbClearResult(result_set)
```

```{r}
#| code-fold: true
p <- jobs |>
    filter(customer_id %in% top20_customers_total$customer_id) |>
    group_by(customer_id) |>
    summarise(n_jobs = n_distinct(job_id)) |>
    arrange(desc(n_jobs)) |>
    ggplot(aes(
        x = fct_reorder(customer_id, -n_jobs),
        y = n_jobs,
        text = paste("Customer:", customer_id, "\nNumber of Jobs:", n_jobs)
    )) +
    geom_col(fill = "#445162") +
    labs(
        title = "Number of jobs by Customer",
        x = "Customer",
        y = "Number of Jobs"
    ) +
    theme_minimal() +
    theme(
        axis.text.x = element_text(angle = 55, hjust = 1)
    )

ggplotly(p, tooltip = "text")
```
### Number of jobs by month of due date
```{r}
#| code-fold: true
p <- jobs |>
    filter(customer_id %in% top20_customers_total$customer_id) |>
    mutate(
        month = month(production_due_date),
        year = year(production_due_date),
        month_year = make_date(year, month)
    ) |>
    group_by(month_year) |>
    summarise(n_jobs = n_distinct(job_id)) |>
    arrange(desc(n_jobs)) |>
    ggplot(aes(
        x = month_year,
        y = n_jobs,
        text = paste(
            "Month:", month_year,
            "\nNumber of jobs:", n_jobs
        )
    )) +
    geom_col(fill = "#445162") +
    labs(
        title = "Number of jobs by Due-Date",
        x = "Date",
        y = "Number of jobs"
    ) +
    theme_minimal() +
    theme(
        axis.text.x = element_text(angle = 45, hjust = 1)
    )

ggplotly(p, tooltip = "text")
```









