---
title: "seasonal trends"
output: html_notebook
---

install libraries
```{r}
shh <- suppressMessages
shh(library(DBI))
shh(library(RPostgres))
shh(library(tidyverse))
shh(library(lubridate))
shh(library(plotly))
```

set up connection to postgres
```{r}
connection <- dbConnect(
  Postgres(),
  dbname = "metalworkingsolutions",
  host = "localhost",
  port = 5432,
  user = "postgres",
  password = rstudioapi::askForPassword("Database Password")
)
```

create query for top 20 customers for seasonal evaluation
```{r}
query <- "
SELECT 
    omp_customer_organization_id,
    omp_sales_order_id,
    omp_customer_po,
    omp_order_date,
    uomp_promise_date,
    omp_payment_term_id,
    omp_order_total_base
FROM sales_orders;
"
result_set <- dbSendQuery(connection, query)
```

create tibble for customers
```{r}
customers <- dbFetch(result_set)
dbClearResult(result_set)
```

```{r}
customers
```


### get top 20 customers by generated revenue for 2023, 2024, and both years combined

2023
```{r}
top20_customers_2023 <- customers |> 
  select(omp_order_date, omp_customer_organization_id, omp_order_total_base) |> 
  group_by(omp_customer_organization_id) |> 
  mutate(omp_order_date = year(omp_order_date)) |> 
  filter(omp_order_date %in% 2023) |> 
  summarise(generated_revenue_2023 = sum(omp_order_total_base)) |> 
  rename(customer_id = omp_customer_organization_id) |> 
  arrange(desc(generated_revenue_2023)) |> 
  select(customer_id) |> 
  head(20)
```

2024
```{r}
top20_customers_2024 <- customers |> 
  select(omp_order_date, omp_customer_organization_id, omp_order_total_base) |> 
  group_by(omp_customer_organization_id) |> 
  mutate(omp_order_date = year(omp_order_date)) |> 
  filter(omp_order_date %in% 2024) |> 
  summarise(generated_revenue_2023 = sum(omp_order_total_base)) |> 
  rename(customer_id = omp_customer_organization_id) |> 
  arrange(desc(generated_revenue_2023)) |> 
  select(customer_id) |> 
  head(20)
```

2023 & 2024
```{r}
top20_customers_total <- customers |> 
  select(omp_order_date, omp_customer_organization_id, omp_order_total_base) |> 
  group_by(omp_customer_organization_id) |> 
  mutate(omp_order_date = year(omp_order_date)) |> 
  summarise(generated_revenue_2023 = sum(omp_order_total_base)) |> 
  rename(customer_id = omp_customer_organization_id) |> 
  arrange(desc(generated_revenue_2023)) |> 
  select(customer_id) |> 
  head(20)
```


## monthly change by customer, 2023
Using plotly for interactivity. 
```{r, out.width='100%'}
p <- customers |> 
  select(omp_order_date, omp_customer_organization_id, omp_order_total_base) |> 
  mutate(month = month(omp_order_date), year = year(omp_order_date)) |> 
  rename(customer_id = omp_customer_organization_id, generated_revenue = omp_order_total_base) |> 
  filter(year == 2023, customer_id %in% top20_customers_2023$customer_id) |> 
  group_by(customer_id, month) |> 
  summarise(generated_revenue = sum(generated_revenue), .groups = 'drop') |> 
  mutate(customer_id = factor(customer_id, levels = top20_customers_2023$customer_id)) |> 
  ggplot(aes(x=month, y=generated_revenue, color=customer_id)) +
  geom_line() +
  scale_x_continuous(
    breaks = seq_along(month.name),
    labels = month.name 
  ) +
  labs(
    title = "Change in Customer Generated Revenue 2023",
    x = "Month",
    y = "Revenue ($)"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 55, hjust = 1)
  ) 
ggplotly(p)
```

ggplot version for displaying inline better.
```{r, fig.width=10}
customers |> 
  select(omp_order_date, omp_customer_organization_id, omp_order_total_base) |> 
  mutate(month = month(omp_order_date), year = year(omp_order_date)) |> 
  rename(customer_id = omp_customer_organization_id, generated_revenue = omp_order_total_base) |> 
  filter(year == 2023, customer_id %in% top20_customers_2023$customer_id) |> 
  group_by(customer_id, month) |> 
  summarise(generated_revenue = sum(generated_revenue), .groups = 'drop') |> 
  mutate(customer_id = factor(customer_id, levels = top20_customers_2023$customer_id)) |> 
  ggplot(aes(x=month, y=generated_revenue, color=customer_id)) +
  geom_line() +
  scale_x_continuous(
    breaks = seq_along(month.name),
    labels = month.name 
  ) +
  labs(
    title = "Change in Customer Generated Revenue 2023",
    x = "Month",
    y = "Revenue ($)"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 55, hjust = 1)
  ) 
```

## monthly change by customer, 2024
Using plotly for interactivity 
```{r}
p <- customers |> 
  select(omp_order_date, omp_customer_organization_id, omp_order_total_base) |> 
  mutate(month = month(omp_order_date), year = year(omp_order_date)) |> 
  rename(customer_id = omp_customer_organization_id, generated_revenue = omp_order_total_base) |> 
  filter(year == 2024, customer_id %in% top20_customers_2024$customer_id) |> 
  group_by(customer_id, month) |> 
  summarise(generated_revenue = sum(generated_revenue), .groups = 'drop') |> 
  mutate(customer_id = factor(customer_id, levels = top20_customers_2024$customer_id)) |> 
  ggplot(aes(x=month, y=generated_revenue, color=customer_id)) +
  geom_line() +
  scale_x_continuous(
    breaks = seq_along(month.name),
    labels = month.name 
  ) +
  labs(
    title = "Change in Customer Generated Revenue 2024",
    x = "Month",
    y = "Revenue ($)"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 55, hjust = 1)
  ) 
ggplotly(p)
```

using ggplot for better inline visibility
```{r, fig.width=10}
customers |> 
  select(omp_order_date, omp_customer_organization_id, omp_order_total_base) |> 
  mutate(month = month(omp_order_date), year = year(omp_order_date)) |> 
  rename(customer_id = omp_customer_organization_id, generated_revenue = omp_order_total_base) |> 
  filter(year == 2024, customer_id %in% top20_customers_2024$customer_id) |> 
  group_by(customer_id, month) |> 
  summarise(generated_revenue = sum(generated_revenue), .groups = 'drop') |> 
  mutate(customer_id = factor(customer_id, levels = top20_customers_2024$customer_id)) |> 
  ggplot(aes(x=month, y=generated_revenue, color=customer_id)) +
  geom_line() +
  scale_x_continuous(
    breaks = seq_along(month.name),
    labels = month.name 
  ) +
  labs(
    title = "Change in Customer Generated Revenue 2024",
    x = "Month",
    y = "Revenue ($)"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 55, hjust = 1)
  ) 

```

2023 + 2024
```{r}
p <- customers |> 
  select(omp_order_date, omp_customer_organization_id, omp_order_total_base) |> 
  mutate(month = month(omp_order_date), 
         year = year(omp_order_date),
         month_year = make_date(year, month)) |> # Create a date object for proper ordering
  rename(customer_id = omp_customer_organization_id, generated_revenue = omp_order_total_base) |> 
  filter(customer_id %in% top20_customers_total$customer_id) |> 
  group_by(customer_id, month_year) |> # Group by month_year
  summarise(generated_revenue = sum(generated_revenue), .groups = 'drop') |> 
  mutate(customer_id = factor(customer_id, levels = top20_customers_total$customer_id)) |> 
  ggplot(aes(x = month_year, y = generated_revenue, color = customer_id)) +
  geom_line() +
  scale_x_date(date_breaks = "1 month", date_labels = "%b %Y") + # Format x-axis
  labs(
    title = "Change in Customer Generated Revenue 2023 + 2024",
    x = "Month",
    y = "Revenue ($)"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 55, hjust = 1)
  )

ggplotly(p)


```

ggplot for inline visibility
```{r, fig.width=10}
customers |> 
  select(omp_order_date, omp_customer_organization_id, omp_order_total_base) |> 
  mutate(month = month(omp_order_date), 
         year = year(omp_order_date),
         month_year = make_date(year, month)) |> # Create a date object for proper ordering
  rename(customer_id = omp_customer_organization_id, generated_revenue = omp_order_total_base) |> 
  filter(customer_id %in% top20_customers_total$customer_id) |> 
  group_by(customer_id, month_year) |> # Group by month_year
  summarise(generated_revenue = sum(generated_revenue), .groups = 'drop') |> 
  mutate(customer_id = factor(customer_id, levels = top20_customers_total$customer_id)) |> 
  ggplot(aes(x = month_year, y = generated_revenue, color = customer_id)) +
  geom_line() +
  scale_x_date(date_breaks = "1 month", date_labels = "%b %Y") + # Format x-axis
  labs(
    title = "Change in Customer Generated Revenue 2023 + 2024",
    x = "Month",
    y = "Revenue ($)"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 55, hjust = 1)
  )
```






shipments missed / later than promise date seasonally 












